---
title: "How to convert casual users of the bike-share service to annual members"
author: "Ryo Hamada"
date: "2025-04-14"
output: html_document
---

## Goal of this analysis

We will dive into the data set of a fictional bike-share company in Chicago.
This data set is about each user's usage history of the service.

We mainly have two purposes.

### 1. Analyze the difference between casual users and annual members.

The company provide users two ways of using the service. Casual user plan, which is pay-as-you-go, and Annual plan.
(From here, we call users with a casual plan customers, and those with an annual plan subscribers.)

We will discover how users are different between the two plans.

### 2. Find the solution to convert customers to subscribers

In the situation that the annual plan is more profitable for the company, we will analyze and consider the good solution to convert existing customers to subscribers.

## Data we used

We have used the data from [here](https://divvybikes-marketing-staging.lyft.net/system-data), which is in public under [this](https://divvybikes.com/data-license-agreement) license.

We have used the 2019/Q1 data.

- data example

```{r}
df2019 <- read.csv('Divvy_trips_2019_Q1.csv')
head(df2019)
```

### cleaning process

```R
df1 <- read.csv('Divvy_trips_2019_Q1.csv')

###### cast data type ######
#  - start_time / end_time -> datetime
#  - tripduration -> float
df1$start_time <- as_datetime(df1$start_time)
df1$end_time <- as_datetime(df1$end_time)
df1$tripduration <- as.numeric(gsub(",", "", df1$tripduration))

###### null check ######
df1[!complete.cases(df1),]

# null row ratio
nrow(df1[!complete.cases(df1),]) / nrow(df1)

# not 100% confident, but probably should remove all rows with missing values
df1 <- na.omit(df1)
df1 <- df1[df1$gender != '',]

###### duplicate check ######
any(duplicated(df1$trip_id))
# -> no duplicates in trip_id

###### outlier check ######
# obviously, there are some mistyped values in tripduration
ggplot(df1) +
  geom_boxplot(aes(y = tripduration))

# it's hard to determine the border between the normal values and anomalies
# for now, I treat values more than the 99th quantile as outliers
df1 <- df1[df1$tripduration <= quantile(df1$tripduration, 0.99),]

table(df1$usertype) # 94% subscriber
table(df1$gender) # 81% male
# -> a little bit biased data set

# treat birthyears before 100 or more years ago (as of 2019) as outliers
df1 <- df1[df1$birthyear > 1919,]

# add a column for the day of the week
df1$day_of_week <- wday(df1$start_time)
```

## Analysis

#### Graph.1

- Average trip duration of each day of the week, for customers and subscribers
  
```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(lubridate)
df1 <- read.csv('Divvy_trips_2019_Q1.csv')
df1$start_time <- as_datetime(df1$start_time)
df1$end_time <- as_datetime(df1$end_time)
df1$tripduration <- as.numeric(gsub(",", "", df1$tripduration))
df1 <- na.omit(df1)
df1 <- df1[df1$gender != '',]
df1 <- df1[df1$tripduration <= quantile(df1$tripduration, 0.99),]
df1 <- df1[df1$birthyear > 1919,]
df1$day_of_week <- wday(df1$start_time)
```

```{r graph1, echo=FALSE}
df1 %>%
  group_by(day_of_week, usertype) %>%
  summarise(avg_tripduration = mean(tripduration), .groups = "drop") %>%
  mutate(day_of_week = factor(day_of_week,
                              levels = 1:7,
                              labels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"))) %>%
  ggplot(aes(x = day_of_week, y = avg_tripduration)) +
  geom_col() +
  facet_wrap(~usertype) +
  labs(x = 'Day of the Week', y = 'Average Trip Duration')
```

From Graph.1, we can see that customers tend to ride more distances than subscribers. This seems counter-intuitive.

**I consider this is because a lot of customers are tourists, or do not ride a bike on daily basis.**

**A lot of them use bikes as a leisure, not just as a means of transportation.**

The fact that the trip duration on Saturdays and Sundays are longer than other days of the week supports this hypothesis, too.

**On the other hand, subscribers tend to ride almost the same distance, throughout the week. This is probably because it is necessary for them to ride a bike in their life.**

They ride a bike less as a leisure, than customers.


#### Graph.2

- The distributions of birth years, for customers and subscribers
  
```{r graph2, echo=FALSE}
ggplot(df1, aes(x = usertype, y = birthyear)) +
  geom_boxplot(aes(color = gender))
```

From graph.2, we can see that subscribers are older than customers. Subscribers are around 35~45 years old, and customers are around 25~35.

We can come up with various reasons for this.

**Those with the age of 35~45 might have more money to pay for the annual plan, and less physical endurance than younger people.**

**In addition, there might be more office workers among those with the age of 35~45 compared to younger generations.**


## Conclusion: how to convert customers to subscribers

From the analysis above, we could infer that subscribers are 

- older(around 35 to 45) than customers, and
- need to ride a bike with a certain amount of time every day.

Therefore, **the company should display ads or launch campaigns for those who are at the age of about 35 to 45, and live in Chicago, and might need a bike on regular basis, such as office workers.**

---

With the data set we have explored, I could not dive into deeper.

Especially, I could not clarify exactly what kind of people need to ride a bike on regular basis every day.

If more data sets that can be merged with the data above are provided, I would try to solve it.

